---
title: "MusicDatabase"
author: Collin Smith, Conor Burrows, Jake Luedtke, Nicholas Melon
date: "January 25, 2017"
output: html_document
---
#Overview
  
In our group our main focus is to determine what genre is the most popular in each decade starting with 1900 and going up to 2016. We also are going to look at which genres are more popular in eruope compared to the U.S and if the words in a song title have a effect on how popular a song is. We are going to use music brainz to get the genre of a song, it's mood quality, and tempo. With this information we will be able to see the trends better in preffered music choice for each decade. In the end we may use the database to make a simple song finder based on your taste in genre, tempo, and mood.

## First Steps: Getting data and Doing stuff

```{r}
library(stringr)
#Get dataset and look at name and title
allsongs <- read.csv("tsort-chart-2-6-0013.csv")
#nsongs <- allsongs[,1:2]
rsongs <- allsongs[,1:3]
justsongs<- rsongs[(rsongs$type=="song"),]
justalbums <- rsongs[(rsongs$type=="album"),]
#Set as characters
csongs <- as.character(justsongs$name)
calbums <- as.character(justalbums$name)
#Make sure song names are in correct format
csongs <- tolower(csongs)
csongs <- str_extract_all(csongs, boundary("word"))
calbums <- tolower(calbums)
calbums <- str_extract_all(calbums, boundary("word"))


#Empty list for words
words <- list()
#Function looks at song titles, splits them up into words, adds them to a list and keeps track of number 
dtf <- data.frame()
for (x in 1:length(csongs)){
  title <- csongs[[x]]
  for (y in 1:length(title)){
    word <- title[[y]]
    if(exists(word, words)){
      words[[word]] <- words[[word]] + 1
    }else{
      words[[word]] <- 1
    }
  }
}
worda <- list()
#Function looks at song titles, splits them up into words, adds them to a list and keeps track of number 
dtfa <- data.frame()
for (x in 1:length(calbums)){
  titlea <- calbums[[x]]
  for (y in 1:length(titlea)){
    worda <- title[[y]]
    if(exists(aword, worda)){
      worda[[aword]] <- worda[[aword]] + 1
    }else{
      worda[[aword]] <- 1
    }
  }
}

#Formatting Data and saving
mdf <- as.data.frame(words)
gmdf <- t(mdf)
colnames(gmdf) <- c("Instances")
write.table(gmdf, file = "songvalues.csv")

mdfa <- as.data.frame(worda)
gmdfa <- t(mdfa)
colnames(gmdfa) <- c("Instances")
write.table(gmdfa, file = "albumvalues.csv")
```

## Various Data Cleaning, Adding Decade Group Variable

```{r}
# Read the updated CSV
chart <- read.csv("chart.csv")

# Cleaning year--replace unknowns with NAs, convert to numeric
chart$year[which(as.character(chart$year) == "unknown")] <- NA
chart$year <- as.numeric(as.character(chart$year))

# Get decade from year
findDecade <- function(year) {
  after1900 <- year-1900
  if (is.na(year)) {
    decade <- NA # Account for NAs
  } else if (after1900 >= 100) {
    decade <- 2000 # Assign 2000 to anything 21st century
  } else {
    tens <- after1900 %/% 10
    decade <- 1900 + 10*tens
  }
  return(decade)
}

decades <- unlist(lapply(chart$year, findDecade))
chart$decade <- decades

# update csv
write.csv(chart, "chart.csv")
```
## GraceNote and Metadata Aggregation
### Required Libraries:
**PYGN:** https://github.com/cweichen/pygn
**TQDM:** https://pypi.python.org/pypi/tqdm

Running GNAW, our unofficial GraceNote API Wrapper, requires you to authenticate via Pygn. If you have a GraceNote developer account, this can be done using your "clientID" from your application page and running pygn.register(clientID). Please note that Gracenote will throttle you by throwing exceptions if you attempt to collect data too many times. However, GNAW takes this into account and lets you run the program on the same file to generate a new output file with more complete data.

```{python}
from __future__ import print_function
from tqdm import tqdm
import pygn, sys, json, csv

# Logins
clientID = '1441931071-E781596CC5CC4E1EBD05775BCC16A317'
userID = '49918418169816731-A8CD882F481B3E8C18D613F8A18E9DF4'

clientID2 = '709371742-00E7F1DE5F12945112375208FDFFFD5B'
userID2 = '27681944132852331-1A932F0545D3D2A22080194840BC047E'


# Init lists
music = []
meta = []

# Open the csv
with open('top5000songs.csv') as songcsv:
    songreader = csv.reader(songcsv, delimiter = ',')
    for row in songreader:
        music.append(row)

with open('metadata5000d.txt', 'r+') as metafile:
    for row in metafile:
        meta.append(row)

index = 0
# Main Loop
with open('metadata5000e.txt', 'a+') as fileout:
    for item in tqdm(music):
        try:
            if meta[index] == "None\n":
                temp = pygn.search(clientID=clientID, userID=userID, artist=item[1], track=item[2])
                fileout.write("%s\n" % temp)
                index += 1
            else:
                temp = meta[index]
                fileout.write(temp)
                index += 1
        except Exception:
            try:
                temp = pygn.search(clientID=clientID2, userID=userID2, artist=item[1], track=item[2])
                fileout.write("%s\n" % temp)
                index += 1
                continue
            except Exception:
                fileout.write("%s\n" % "None")
                index += 1
                continue

```


